Dataflow-Guided Filtering for Efficient and Adjustable Run-Time Monitoring

### 2014-05-28 drop.tex
Filled out extensions section.
Various edits.

The goal is to be able to perform as much monitoring as possible within a specified overhead budget.
The high-level approach is to keep track of monitoring overheads at run-time
and drop monitoring operations when the overhead exceed the specified budget.
A block diagram of the architecture we will present in this section is shown in
Figure~\ref{fig:drop.overview}.  In order to do limit the overheads of
monitoring, we need to first be able to track the gap between the specified
overhead budget and the monitoring overhead seen at run-time. We refer to this
as slack and discuss how we measure it in Section~\ref{sec:drop.slack}. 
Dropping monitoring operations can cause false positives which we prevent by
extending the dataflow engine to track and filter out dropped monitoring flows (see
Section~\ref{sec:drop.false_positives}).
Given that we know how much slack we have for monitoring and can drop
monitoring operations without false positives, we must decide when and which
monitoring operations to drop.  There exists some interesting trade-offs in the
policy for this dropping decision which we discuss in
Section~\ref{sec:drop.policies}.

@@

% Flag propagation rules for BC
\begin{table}[tb]
  \begin{center}
    \begin{small}
    \input{tabs/propagation_rules}
    \end{small}
    \caption{Flag propagation rules for array bounds check. A dash indicates that no
    propagation is done while 0 or 1 indicates that the destination flag is set
    to that value.}
    \label{tab:drop.propagation_rules}
    \vspace{-0.2in}
  \end{center}
\end{table}

We can track these invalid flags by extending the dataflow engine presented in
Section~\ref{sec:filter} to keep track of an additional bit for each flag. That
is, we use the dataflow engine to track 2-bit flags in order to keep track of
both clean/dirty information as well as valid/invalid information.
Similar to filtering for clean flags, we can also filter out
monitoring corresponding to invalid metadata which would be useless to perform.
However, it is important that we propagate the
invalidation flag in these circumstances so that the entire flow is marked as invalid. 
In the typical case, the propagation rule for this is
that if any of the source operands is marked as invalid, then the destination
operand is marked as invalid. This is slightly different than the
policy for clean flags which propagates a clean flag only if both
sources are marked as clean. Table~\ref{tab:drop.propagation_rules} shows the
propagation rules that are used for array bounds check. However, we note that
the propagation rules can be customized depending on the monitor by
reconfiguring the filter decision table.

@@

The simplest policy of when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be also dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results not used for any monitoring checks.

@@

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-event dropping} and we refer to the previous policy of waiting until
the overhead budget is reached as \emph{all-event dropping}.
Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

@@

\textbf{[We probably need more details of how min-event dropping is
implemented. How is analysis done? How is the information conveyed to
run-time?]}

>>>>

The goal is to be able to perform as much monitoring as possible within a specified overhead budget.
The high-level approach is to keep track of monitoring overheads at run-time
and drop monitoring operations when the overhead exceed the specified budget.
A block diagram of the architecture we will present in this section is shown in
Figure~\ref{fig:drop.overview}.  In order to limit the overheads of
monitoring, we need to first be able to track the gap between the specified
overhead budget and the monitoring overhead seen at run-time. We refer to this
as slack and discuss how we measure it in Section~\ref{sec:drop.slack}. 
Dropping monitoring operations can cause false positives which we prevent by
extending the dataflow engine to track and filter out dropped monitoring flows (see
Section~\ref{sec:drop.false_positives}).
Given that we know how much slack we have for monitoring and can drop
monitoring operations without false positives, we must decide when and which
monitoring operations to drop.  There exists some interesting trade-offs in the
policy for this dropping decision which we discuss in
Section~\ref{sec:drop.policies}.

@@

% % Flag propagation rules for BC
% \begin{table}[tb]
%   \begin{center}
%     \begin{small}
%     \input{tabs/propagation_rules}
%     \end{small}
%     \caption{Flag propagation rules for array bounds check. A dash indicates that no
%     propagation is done while a 1 indicates that the destination flag is set.}
%     \label{tab:drop.propagation_rules}
%     \vspace{-0.2in}
%   \end{center}
% \end{table}

We can track these invalid flags by extending the dataflow engine presented in
Section~\ref{sec:filter} to keep track of an additional bit for each flag. That
is, we use the dataflow engine to track 2-bit flags in order to keep track of
both clean/dirty information as well as valid/invalid information.
Similar to filtering for clean flags, we can also filter out
monitoring operations corresponding to invalid metadata which would be useless to perform.
However, it is important that we propagate the
invalidation flag in these circumstances so that the entire flow is marked as invalid. 
In the typical case, the propagation rule used is
that if any of the source operands is marked as invalid, then the destination
operand is marked as invalid. This is slightly different than the
policy for clean flags which propagates a clean flag only if both
sources are marked as clean. 
% Table~\ref{tab:drop.propagation_rules} shows the
% propagation rules that are used for array bounds check. 
We note that
the propagation rules can be customized depending on the monitor by
reconfiguring the filter decision table.

@@

The simplest policy for when to drop monitoring events is to drop all monitoring events as
long as slack is less than or equal to zero.  However, this can result in
wasted work. For example, consider the metadata dependence graph shown in
Figure~\ref{fig:drop.dataflow_graph}. Here, an edge from node {\tt A} to node
{\tt B} represents that if event {\tt A} is dropped, then due to its
invalidated metadata, it will cause event {\tt B} to be also dropped. Square
nodes indicate events where monitoring checks are performed. In the
example, suppose that event {\tt E} is meant to perform a check operation but is dropped.
In this situation, the
monitoring operations that were done for events {\tt C} and {\tt D} were wasted
since their results were not used for any monitoring checks.

@@

One method to eliminate this wasted work is to only make dropping decisions at
the root of these metadata flows. That is, we will decide to either monitor or
not monitor an entire metadata flow. We refer to this dropping decision policy
as \emph{source-event dropping} and we refer to the previous policy of 
dropping any event as \emph{all-event dropping}.
Figures~\ref{fig:drop.all_drop} and \ref{fig:drop.source_drop} shows a
comparison of where dropping decisions are made for these two policies. Source
dropping will
result in no wasted work. However, because of the coarser-grained decision, it
may be more difficult to closely match overheads. Source dropping is expected
to work well when there are a large number of independent metadata flows.

@@

### 2014-05-28 extensions.tex
Various edits.

On a dropped store instruction, the destination memory location is marked as
invalid. On load instructions, we can filter them out if the source memory
location has been invalidated. In this case, we do not have to perform any
further operation since on load instructions, UMC only performs a check
operation.

@@

Dynamic information flow tracking (DIFT) \cite{dift-asplos04} is a security
monitoring scheme that seeks to detect when information from untrusted sources
is used to affect the control flow.  In its simplest form, DIFT keeps a 1-bit
metadata tag for each memory location and each register, indicating tainted or
untainted. When data is read from an untrusted source, it is marked as tainted.
As this data is used for other operations, the results of these operations are
also marked as tainted. On an indirect jump, the taint bit of the register used
is checked. If the register is found to be tainted, then an error is detected.
We focus on a more heavy-weight approach that also allows some information of
where flows originate from by using a multi-bit (32-bits in our implementation)
metadata tag. This can be used to assist debugging in the case of an error. In
this case, when data is marked from an untrusted source, the software specifies
an identifier to set the metadata tag. This multi-bit tag is propagated on
load, store, and ALU operations. If two different tags are used as sources to
an ALU operation, then a new identifier is calculated based on the sources.
These monitoring operations are summarized in
Table~\ref{tab:extensions.monitor}.

@@

Again, the invalid flags for DIFT operate similar to their operation for BC
(Table~\ref{tab:extensions.drop}. In addition to operations for invalidating
destination operands for ALU, load, and store instructions, the destination
memory location is also invalidated when set tag operations dropped. Indirect
jump events are filtered out if the source register is marked as invalid.
However, no further metadata needs to be marked as invalid in this case.

@@

We do note some limitations to the proposed approach. Previous work
\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
overheads has also filtered out redundant updates where the source metadata
matches the destination metadata. Since we do not read the actual metadata, our
architecture is not able to identify these redundant updates. However, by using
only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
In addition, we note that filtering use clean flags works well when a large
portion of instructions are not relevant to the monitoring scheme (e.g., array
bounds check only cares about array pointers). However, if most instruction
have dirty metadata, then very little filtering will be possible.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination.  For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a clean and invalid flag with a single piece.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as clean may indicate that the entire data structure's metadata
is clean.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support these more complex monitoring schemes.

>>>>

On a dropped store instruction, the destination memory location is marked as
invalid. On load instructions, we can filter them out if the source memory
location has been invalidated. In this case, we do not have to perform any
further operation since on load instructions since UMC only performs a check
operation for loads.

@@

Dynamic information flow tracking (DIFT) \cite{dift-asplos04} is a security
monitoring scheme that seeks to detect when information from untrusted sources
is used to affect the control flow.  In its simplest form, DIFT keeps a 1-bit
metadata tag for each memory location and each register, indicating tainted or
untainted. When data is read from an untrusted source, it is marked as tainted.
As this data is used for other operations, the results of these operations are
also marked as tainted. On an indirect jump, the taint bit of the register used
is checked. If the register is found to be tainted, then an error is detected.
We focus on a more heavy-weight approach that also tracks some information of
where flows originate from by using a multi-bit (32-bits in our implementation)
metadata tag. This can be used to assist debugging in the case of an error. In
this case, when data is marked from an untrusted source, the software specifies
an identifier to set the metadata tag. This multi-bit tag is propagated on
load, store, and ALU operations. If two different tags are used as sources to
an ALU operation, then a new identifier is calculated based on the sources.
These monitoring operations are summarized in
Table~\ref{tab:extensions.monitor}.

@@

Again, the invalid flags for DIFT operate similar to their operation for BC
(Table~\ref{tab:extensions.drop}). In addition to operations for invalidating
destination operands for ALU, load, and store instructions, the destination
memory location is also invalidated when set tag operations dropped. Indirect
jump events are filtered out if the source register is marked as invalid.
However, no further metadata needs to be marked as invalid in this case.

@@

We do note some limitations to the proposed approach. Previous work
\cite{fade-hpca14} that has looked at filtering in order to reduce monitoring
overheads has also filtered out redundant updates where the source metadata
matches the destination metadata. Since we do not read the actual metadata, our
architecture is not able to identify these redundant updates. However, by using
only 2-bit flags, the bandwidth need of our dataflow engine is reduced.  
In addition, we note that filtering using clean flags works well when a large
portion of instructions are not relevant to the monitoring scheme (e.g., array
bounds check only cares about array pointers). However, if most instruction
have dirty metadata, then very little filtering will be possible.

@@

Another limitation is that the dataflow engine, as described, only marks one
flag location on each monitoring event. Similarly, the dataflow engine can only
read in two sets of flags in order to determine whether an event can be
filtered or not. This was done because we found that for many monitoring
schemes, the monitoring operations corresponded to program instructions, which
read up to two operands and write to one destination.  For monitoring schemes
which update multiple pieces of metadata on a monitoring event, it may be
adequate to associate a clean and invalid flag with a single piece of metadata.  For
example, a scheme may use a data structure that includes several pieces of
related metadata. On a drop, one of these pieces could be marked as invalid in
order to indicate that the entire structure is invalid.  Similarly, marking one
specified piece as clean may indicate that the entire data structure's metadata
is clean.  Alternatively, the read and write capabilities of the dataflow
engine could be expanded to support more complex monitoring schemes.

### 2014-05-28 introduction.tex
Various edits.

Thus, the main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
  \item We enable a trade-off between monitoring coverage and overheads by extending the dataflow tracking engine to keep track of dropped monitoring flows.
  \item We explore the design space and trade-offs in dropping policies for performing partial monitoring.
\end{itemize}

>>>>

Thus, the main contributions of this paper are summarized as follows.
\begin{itemize}
  \item We use a dataflow tracking engine to efficiently filter out monitoring events in order to reduce overheads.
  \item We present an architecture that enables adjustable monitoring, creating a trade-off between monitoring coverage and overheads.
  \item We show how the dataflow tracking engine can be used to prevent false positives for adjustable monitoring.
  \item We explore the design space and trade-offs in dropping policies for performing adjustable monitoring.
\end{itemize}

### 2014-05-28 related.tex
Various edits.

% Filtering
FADE \cite{fade-hpca14} provides a more general filtering engine. It
filters out redundant updates and clean checks by reading metadata values. Our
dataflow engine does not require reading the actual metadata values which may
be large. We do not handle filtering redundant updates, but we do filter out
clean checks as well as clean (i.e., uninitialized) propagations. We are able
to do this because the dataflow engine can also propagate the uninitialized
information. In addition, our dataflow engine can be used to provide adjustable
overheads without false positives which is not a capability described by FADE.

>>>>

% Filtering
FADE \cite{fade-hpca14} presents a general filtering engine in order to reduce
the overheads of monitoring. It filters out redundant updates and clean checks
by reading metadata values. We do not handle filtering redundant updates, but
we do filter out clean checks as well as clean (i.e., uninitialized)
propagations.  We are able to do this because the dataflow engine can also
propagate the clean information.  In addition, our dataflow engine only
requires reading and writing 2-bit flags rather than the entire metadata.  As
we have shown, our dataflow engine also enables adjustable overheads without
false positives which is not a capability described by FADE.

### 2014-05-28 tabs/monitor.tex
Various edits.

\multirow{5}{*}{DIFT}  
 & Software set tag & Set tag to a specified value based on software command \\ \cline{2-3}
 & ALU & Output tag is the OR of the input tags \\ \cline{2-3}
 & Load & Set register tag as tag of loaded data \\ \cline{2-3}
 & Store & Set memory tag as register tag \\ \cline{2-3}
% & Software clear tag & Clear tag & Update \\ \cline{2-4}
 & Indirect jump & Check tag \\ 
 \hline

>>>>

\multirow{5}{*}{DIFT}  
 & Software set tag & Set tag to a specified identifier value based on software command \\ \cline{2-3}
 & ALU & If input tags are different, output tag is a new identifier. Otherwise, output tag is the input tag. \\ \cline{2-3}
 & Load & Set register tag as tag of loaded data \\ \cline{2-3}
 & Store & Set memory tag as register tag \\ \cline{2-3}
% & Software clear tag & Clear tag & Update \\ \cline{2-4}
 & Indirect jump & Check tag \\ 
 \hline
